---
author: "PrachiJhamb"
date: "3/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.height=4)
necessaryPackages <- c("rvest","tidyverse","dplyr","stringr","ggplot2")
new.packages <- necessaryPackages[
              !(necessaryPackages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(necessaryPackages, require, character.only = TRUE)
```
**Task: Analyse the Athens Real Estate Market**
**Run a housing search.Open the zillow website in your browser and play around with the search function.**

```{r}
library(rvest)
library(tidyverse)
library(stringr)

page1 <- read_html("https://www.zillow.com/athens-ga/2-_beds/?searchQueryState=%7B%22usersSearchTerm%22%3A%22Athens%2C%20GA%22%2C%22mapBounds%22%3A%7B%22west%22%3A-83.63818270019532%2C%22east%22%3A-83.16439729980469%2C%22south%22%3A33.76266117514294%2C%22north%22%3A34.1533615364639%7D%2C%22mapZoom%22%3A11%2C%22regionSelection%22%3A%5B%7B%22regionId%22%3A23534%2C%22regionType%22%3A6%7D%5D%2C%22isMapVisible%22%3Atrue%2C%22filterState%22%3A%7B%22beds%22%3A%7B%22min%22%3A2%7D%2C%22ah%22%3A%7B%22value%22%3Atrue%7D%2C%22sort%22%3A%7B%22value%22%3A%22globalrelevanceex%22%7D%2C%22auc%22%3A%7B%22value%22%3Atrue%7D%2C%22fsbo%22%3A%7B%22value%22%3Atrue%7D%7D%2C%22isListVisible%22%3Atrue%7D")

page2 <- read_html("https://www.zillow.com/athens-ga/2-_beds/2_p/?searchQueryState=%7B%22usersSearchTerm%22%3A%22Athens%2C%20GA%22%2C%22mapBounds%22%3A%7B%22west%22%3A-83.63818270019532%2C%22east%22%3A-83.16439729980469%2C%22south%22%3A33.76266117514294%2C%22north%22%3A34.1533615364639%7D%2C%22mapZoom%22%3A11%2C%22regionSelection%22%3A%5B%7B%22regionId%22%3A23534%2C%22regionType%22%3A6%7D%5D%2C%22isMapVisible%22%3Atrue%2C%22filterState%22%3A%7B%22beds%22%3A%7B%22min%22%3A2%7D%2C%22ah%22%3A%7B%22value%22%3Atrue%7D%2C%22sort%22%3A%7B%22value%22%3A%22globalrelevanceex%22%7D%2C%22auc%22%3A%7B%22value%22%3Atrue%7D%2C%22fsbo%22%3A%7B%22value%22%3Atrue%7D%7D%2C%22isListVisible%22%3Atrue%2C%22pagination%22%3A%7B%22currentPage%22%3A2%7D%7D")

page3 <- read_html("https://www.zillow.com/athens-ga/2-_beds/3_p/?searchQueryState=%7B%22usersSearchTerm%22%3A%22Athens%2C%20GA%22%2C%22mapBounds%22%3A%7B%22west%22%3A-83.63818270019532%2C%22east%22%3A-83.16439729980469%2C%22south%22%3A33.76266117514294%2C%22north%22%3A34.1533615364639%7D%2C%22mapZoom%22%3A11%2C%22regionSelection%22%3A%5B%7B%22regionId%22%3A23534%2C%22regionType%22%3A6%7D%5D%2C%22isMapVisible%22%3Atrue%2C%22filterState%22%3A%7B%22beds%22%3A%7B%22min%22%3A2%7D%2C%22ah%22%3A%7B%22value%22%3Atrue%7D%2C%22sort%22%3A%7B%22value%22%3A%22globalrelevanceex%22%7D%2C%22auc%22%3A%7B%22value%22%3Atrue%7D%2C%22fsbo%22%3A%7B%22value%22%3Atrue%7D%7D%2C%22isListVisible%22%3Atrue%2C%22pagination%22%3A%7B%22currentPage%22%3A3%7D%7D")

all_pages <- list(page1, page2, page3)
str(all_pages)  
```
**Scrape the results of this housing search**
I extracted **Price , House details(number of bathrooms, number of bedrooms, square footage,) and Addresses**
```{r}
getPrice <- function(html){ 
  html %>% html_nodes('.list-card-price') %>% html_text() %>% str_trim()
}
getDetails <- function(html){ 
  html %>% html_nodes('.list-card-details') %>% html_text() %>% str_trim()
}
getAddress <- function(html){ 
  html %>% html_nodes('.list-card-addr') %>% html_text() %>% str_trim()
}

rawData = NULL
for (i in 1:3){
  price <- getPrice(all_pages[[i]])
  details <-  getDetails(all_pages[[i]])
  address <-  getAddress(all_pages[[i]])
  rawData <-  rbind(rawData, data.frame(cbind(price, details, address)))
}

```
**Do some cleaning**

```{r}
df1 <- rawData %>%
  mutate(bedrooms = as.integer(str_trim(str_extract(details, "[\\d ]*(?=bds)")))) %>%
  mutate(bathrooms = as.integer(str_trim(str_extract(details, "[\\d ]*(?=ba)")))) %>%
  mutate(sqft = str_trim(str_extract(details, "[\\d ,]*(?=sqft)"))) %>%
  mutate(sqft = as.numeric(str_replace(sqft,",",""))) %>%
  mutate(price = as.numeric(str_replace_all(price,"[^0-9]*","")))
df1 <- na.omit(df1)
#str(df1)
df2 <- as.data.frame(df1)
head(df2)
```
**Here is a  plot of Square footage vs. price**
```{r}
library(ggplot2)
g1 <- ggplot(df2, aes(x=sqft, y=price, size=sqft, color=as.factor(bedrooms))) + 
  geom_point() +
  labs(y = "Price ($)",
       x = "Square footage (squared feet)",
       title = "Square footage vs. Price of houses Athens GA")
g1
```
We see a positive relationship between square footage and price. 


**Histogram showing Number of bathroom in the house**
```{r}
countbaths <- ggplot(data=df2, aes(bathrooms)) + geom_histogram()+
                labs(title="Histogram for Bathrooms") +
                    labs(x="number of bathrooms", y="Count")
countbaths
```
**Regression analysis and goodness of fit**
```{r}
#model 1
linear_model <- lm(price ~ bedrooms + sqft + bathrooms, data = df2)
linear_model$coefficients
summary(linear_model)
```
On average, having an additional bedroom decreases the price by $157541 and this result is significant at the 5% level

On average, having an additional bathroom increases the price by $22539 but this result is not signifcant.
The R-squared for this model is 77 percent. 
**What elements should be considered in this model to increase the fit?**
For a better fitted model, we would need data for variables like the air-pollution, number of nearby schools, recreational centres, hospitals, crime rate etc.

```{r}
#model 2
#non-linear 
nonlinear_model <- lm(log(price)~log(bedrooms)+log(sqft), data=df2)
summary(nonlinear_model)
nonlinear_model$coefficients
```

