---
author: "PrachiJhamb"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval = TRUE, message = FALSE,warning = FALSE,fig.height = 5)
necessaryPackages <- c("foreign","reshape" ,"tidyverse" ,"dplyr" ,"rvest","stringr" ,"ggplot2" ,"stargazer" ,"readr" ,"haven","broom","huxtable","maps")
new.packages <- necessaryPackages[!(necessaryPackages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(necessaryPackages, require, character.only = TRUE)
```

```{r}
setwd("~/Rclass")
df1 <- read.csv("1976-2020-president.csv",1)
str(df1)
summary(df1$candidatevotes)
```


**2. Load the data, and clean it up**
You should aim to make a dataset at the state*year level, with infomation on the party who won in that state and that year.

• Tip 1: you can define the color in that dataset already. Make sure it is defined as.character
• Tip 2: the hardest thing here is to select the winning observations. There are hundreds of ways to do it in R, most of them un-intuitive. My favorite involves group_by then top_n, both from the dplyr
package.
```{r}
library(dplyr)
df1 <- dplyr::rename(df1, party = party_simplified)
winners <- df1 %>% 
  group_by(state,year) %>% 
  top_n(1,candidatevotes)
df2 <- dplyr::select(winners, year, state, state_fips, party, candidate)
df2 <- mutate(df2, color = recode(as.character(party), "DEMOCRAT" = "blue", "REPUBLICAN" = "red", "OTHER" = "white"))
head(df2)
```

**3. Try to make a map just for one year:**
If your data is in the right format, this takes very few commands:
- one command to subset data for just one year
- one command to select the observsations that match state.fips$fips 
- one command to map

```{r}
library(maps)
df3_2020 <- dplyr::filter(df2, year ==2020)
df3_2020 <- df3_2020[match(state.fips$fips,df3_2020$state_fips),]
maps::map('state', col = df3_2020$color, fill = TRUE)
legend("bottomright", legend = c("Democrat", "Republican"), fill = c("blue", "red"), cex = 0.7)
title("2020 Presidential Elections Results", line = 1)
```

**4. Now loop that code and map it over time:**
• If you are using base plots, you will need to define a layout grid
• If you are using ggplot, you can probably do it with facets, but I havent tried.

```{r}
par(mfrow = c(4,3),mar=c(0,0,0,0))
library(dplyr)
# loop over all years
for (i in seq(min(df2$year),max(df2$year),4)) {
  time <- paste("win", i) # create a temporary dataframe for each year
  time <- dplyr::filter(df2, year == i) # select year as defined by 'i' counter 
  time <- time[match(state.fips$fips, time$state_fips),] # match fips
maps::map('state', col = time$color, fill = T) # and plot
title(i, line =1)
}
legend("right", legend = c("Democrat", "Republican","Other"), fill = c("blue", "red","white"), cex = 0.7)
```
**Part 2: Interactive Maps with Leaflet**
1. Get familiar with the leaflet package
```{r}
library(leaflet)
myMap<-leaflet() %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lat = 33.947474,lng = -83.373671,zoom = 12)
myMap
```

**2. Add your own shapefiles**
Go to [data.humdata.org](https://data.humdata.org/dataset/outline-of-camps-sites-of-rohingya-refugees-in-cox-s-bazar-bangladesh) and download the shapefile for Rohingya refugee camps. There are several files, but I suggest using the zipfile ending with “_AL2.zip” (not .kmz file)
Unzip the file and load the .shp into R like this:
```{r}
library(raster)
sp<-shapefile("220130_RRC_Outline_Block_AL2/T220130_RRC_Outline_Block_AL2.shp")
# Projection is necessary for R to place the coordinates correctly
campShapeFile<-spTransform(sp,CRS("+proj=longlat +datum=WGS84 +no_defs"))
head(campShapeFile)
```
• Zoom your map onto South-Eastern Bangladesh (setView(92.14871, 21.18780, zoom = 12)) 
• Add the shapefile to your map with:addpolygons;add some highlights inside addPolygons function ; add labels with another option.

```{r}
library(leaflet)
myMap2<-leaflet() %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  setView(lat = 21.18780,lng = 92.14871,zoom = 12) %>%
  addPolygons(data = campShapeFile,fill = TRUE,stroke = T,weight = 1,highlight=highlightOptions(fillOpacity = 0.7),label = campShapeFile$Block_No)
myMap2
```














